---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

Focus on readr package. 

## Read in some data

absolute path vs. relative paths (always use relative paths)
  * absolute paths are easier, relative path requires you to know the starting point
  * R give you a starting point - your file folder 
  * relative paths is table, anybody downloarding the file can use it

Read in the litters datasets.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df) 
```

## Take a look at the data

Printing in the console.
```{r}
litters_df
```

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```


## Options to read_csv

```{r}
litters_df = read_csv("./data/Fas_litters.csv", skip = 10, col_names = FALSE)
```

```{r}
litters_df = read_csv("./data/Fas_litters.csv", na = c("","NA", 999))
```

check out `?read_csv()` for more information. 

## Other file formats

Read in an excel file.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Read in a SAS file.

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```



